---
title: "Figure 1 Genus Phylum"
author: "Aarthi Ravikrishnan"
date: "2024-02-21"
editor_options: 
  chunk_output_type: console
chunk_output_type: console
output: 
  
  rmdformats::downcute:
    
    downcute_theme: "chaos"
    use_bookdown: true
    self_contained: true
    lightbox: true
    gallery: true
    toc_depth: 5
  pdf_document:
    toc_depth: 5
---

This script does GLM and reports the associations of genus and phylum. 

#Packages  
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE, 
                      collapse = TRUE, warning = FALSE, message = FALSE)

rm(list=ls())
pacman::p_load(tidyverse, janitor,
               knitr,
               tibble,
               here,
               MMUPHin)

```
#Setup parameters
```{r setupparam}
species_thr <- 0.1
genus_thr <- 0.1
hist_filter <- 50
figsfoldername <- paste0("figs", "_", species_thr)
dir.create(file.path(here(figsfoldername)))
resultsfoldername <- paste0("results", "_", species_thr)
dir.create(file.path(here(resultsfoldername)))
editablepptfname <- "figppts"
dir.create(file.path(here(editablepptfname)))
source(here('script','utils_Fig1.R'))
source(here('script','fig_plot.R'))
```
#Genus profiles 

## Reading in profiles
```{r genus}
base_genus_profiles <-
  read.csv(
    here("data", "genus_abundances.csv"),
    encoding =  "UTF-8-BOM") 

cre_genus_profiles <- read.csv(here("data", 'merged_all_genus_abundances.csv')) %>% 
  dplyr::select(-X)

#metadata <- read.csv(here('data','metadata_with_clinical_data_updated_age.csv'))

metadata <- read.csv(here("data", "metadata.csv")) %>% 
  # convert CRE to SG90
  mutate(Cohort = ifelse(Cohort == 'CRE', 'SG90', Cohort))
metadata <- metadata[order(metadata$LibraryID),] 

genus_profiles <- left_join(base_genus_profiles, cre_genus_profiles, by='Index') %>%
  column_to_rownames('Index') %>% 
  dplyr::select(metadata %>% pull(LibraryID)) %>% 
  replace(is.na(.), 0)
```
## Filtering profiles
**Filter criterion**

We employ a two-step filtering criterion:
1. Filtering the profiles by setting abundance < 0.1 as 0 and, 
2. Removing the organisms which are zero across all samples

We then renormalise the profiles to 1.
```{r genusfilter}
genus_thr <- 0.1
genus_profiles_copy <- genus_profiles
genus_profiles[genus_profiles < genus_thr] <- 0
genus_profiles_filtered <-
  genus_profiles[rowSums(genus_profiles) != 0, ]

genus_profiles_prop <-
  genus_profiles_filtered %>% apply(2, function(x)
    x / sum(x))

genus_profiles_prop <-
  genus_profiles_prop[, order(colnames(genus_profiles_prop))]

```
## Batch adjustment
```{r genusbatch}
rownames(metadata) <- metadata$LibraryID
genus_fit_adjust_batch_decade <-
  adjust_batch(
    feature_abd = genus_profiles_prop,
    batch = "Cohort",
    covariates = "AgeG",
    data = metadata
  )
#
genus_profiles_adj_decade <- genus_fit_adjust_batch_decade$feature_abd_adj
genus_profiles_adj_scaled_to_100 <- 100 * genus_profiles_adj_decade
```
## Genus Association
```{r genusasscn}
genus_nonzero_in_samples <- 
  data.frame(nnzorg = colSums(t(genus_profiles_adj_scaled_to_100)!=0))

genus_hist <- ggplot(genus_nonzero_in_samples, aes(x=nnzorg)) +
  geom_histogram(color="darkblue", fill="lightblue") + 
  scale_x_continuous(breaks = scales::pretty_breaks(n=7)) +
  xlab('Number of samples') + 
  ylab('Count of organisms with non-zero abundances') + 
  ggtitle('Genus histogram')

print(genus_hist)

genusinmorethanfiftysamples <-
  genus_nonzero_in_samples %>% filter(nnzorg >= hist_filter)

genus_needed <- genus_profiles_adj_scaled_to_100 %>%
  t() %>% as.data.frame() %>%
  dplyr::select(rownames(genusinmorethanfiftysamples)) %>%
  rownames_to_column('LibraryID')
```

## Read metadata and prepare data
```{r readmetadata}
metadata_with_clinicalparam <- read.csv(here('data', 'metadata_with_imputed_updatedage_allInfo.csv'))

genus_data_for_association <-
  merge(genus_needed, metadata_with_clinicalparam, by = 'LibraryID')

genus_data_for_association <- left_join(genus_data_for_association,
                                        metadata %>% 
                                          dplyr::select('LibraryID',
                                                        'AgeG'),
                                        by='LibraryID')
dim(genus_data_for_association)
dim(metadata)
dim(genus_needed)
genus_list_to_use <-colnames(genus_needed)[grepl('g__',colnames(genus_needed))]
length(genus_list_to_use)
```
## Age association
### GLM
```{r genusglm}
genus_with_age <-
  lapply(genus_data_for_association[, genus_list_to_use], function(x)
    coefficients(summary(
      glm(Age ~ x +  FBGlu +  Tgly + TCHOL + HDL + BMI + Gender,
        data = genus_data_for_association))))

genus_fname <- 'genus_associations_noldl.csv'
genus_results <- param_with_age_correction(genus_list_to_use, genus_with_age,resultsfoldername, genus_fname, genus_data_for_association)

# genus_maaslin2 <- run_maaslin2(genus_data_for_association, 'NONE', 'NONE', 'results_0.1')
# genus_maaslin2_glm <- genus_maaslin2 %>% 
#   dplyr::select(org_list, pval, coef, stderr, padj) %>% 
#   dplyr::rename(betaestimate = coef, stderror = stderr) %>% 
#   # add 95% confint
#   mutate(`2.5%` = betaestimate - 1.96 * stderror, 
#          `97.5%` = betaestimate + 1.96 * stderror)


```
### Preparing data to plot
```{r datapreptoplot}
median_of_genus <- genus_needed %>%
  dplyr::select(rownames(genusinmorethanfiftysamples)) %>% 
  as.matrix() %>% 
  matrixStats::colMedians()

median_of_genus_with_name <- data.frame(rownames(genusinmorethanfiftysamples), median_of_genus)
colnames(median_of_genus_with_name)[1] <- c('org_list')

genus_results$Significant <-
  ifelse(genus_results$padj < 0.05, "FDR < 0.05", "Not Sig")

genus_results_merge <-
  merge(genus_results, median_of_genus_with_name, by = 'org_list')
  #merge(genus_maaslin2_glm, median_of_genus_with_name, by = 'org_list')

genus_results_merge$highlight1 <-
  ifelse(genus_results_merge$median_of_genus > 0 &
           genus_results$padj < 0.05, "2",
         ifelse(genus_results_merge$median_of_genus == 0 &
                  genus_results$padj < 0.05 |
                  genus_results$padj == 0.05 , "1", "0"))

genus_results_merge$org_list <- str_replace_all(genus_results_merge$org_list, 'g__', '')
genus_results_merge$org_list <- str_replace_all(genus_results_merge$org_list, '_', ' ')

write.csv(genus_results_merge, here(resultsfoldername, genus_fname))

```

### Genus boxplot
```{r}
bacteroides_genus <- plotbox(genus_data_for_association, 'AgeG.x', 'g__Bacteroides', 'Bacteroides genus', figsfoldername, 'Fig1D_genus.png',ylimitplot=100,genus_results)
bacteroides_genus
template_to_save(bacteroides_genus, fname = 'Fig1D_genus.png', width=10, height=10)

makeedtiableppt(bacteroides_genus, fname = 'Fig1D_genus.pptx', width = 4, height = 4)
```
# Phylum profiles
## Reading in profiles
```{r phylumtaxaread}
# Reading data
base_phylum_profiles <-
  read.csv(
    here("data", "phylum_abundances.csv"),
    encoding =  "UTF-8-BOM")


cre_phylum_profiles <- read.csv(here("data", 'phylum_cre_profiles.csv')) %>%  dplyr::select(-X)

phylum_profiles <- left_join(base_phylum_profiles, cre_phylum_profiles, by='Index') %>%
  column_to_rownames('Index') %>% 
  dplyr::select(metadata %>% pull(LibraryID)) %>% 
  replace(is.na(.), 0)
```
## Filtering profiles
```{r phylumtaxafilter}
phylum_thr <- 0.1
phylum_profiles_copy <- phylum_profiles
phylum_profiles[phylum_profiles < phylum_thr] <- 0
phylum_profiles_filtered <-
  phylum_profiles[rowSums(phylum_profiles) != 0, ]
phylum_profiles_prop <-
  phylum_profiles_filtered %>% apply(2, function(x)
    x / sum(x))
phylum_profiles_prop <-
  phylum_profiles_prop[, order(colnames(phylum_profiles_prop))]
```
## Batch adjustment
```{r phylumbatchadj}
phylum_fit_adjust_batch_decade <-
  adjust_batch(
    feature_abd = phylum_profiles_prop,
    batch = "Cohort",
    covariates = "AgeG",
    data = metadata
  )
#
phylum_profiles_adj_decade <- phylum_fit_adjust_batch_decade$feature_abd_adj
phylum_profiles_adj_scaled_to_100 <- 100 * phylum_profiles_adj_decade
```
## Phylum Association
```{r phylumasscn}

phylum_nonzero_in_samples <- data.frame(nnzorg = colSums(t(phylum_profiles_adj_scaled_to_100) !=0))

phylum_hist <- ggplot(phylum_nonzero_in_samples, aes(x=nnzorg)) + 
  geom_histogram(color="darkblue", fill="lightblue") +
  scale_x_continuous(breaks = scales::pretty_breaks(n=7)) + 
  xlab('Number of samples') +
  ylab('Count of organisms with non-zero abundances') + 
  ggtitle('phylum histogram')

phyluminmorethanfiftysamples <-
  phylum_nonzero_in_samples %>% filter(nnzorg >= hist_filter)

phylum_needed <- phylum_profiles_adj_scaled_to_100 %>% t() %>% as.data.frame() %>% dplyr::select(rownames(phyluminmorethanfiftysamples)) %>% rownames_to_column('LibraryID')

phylum_data_for_association <-
  merge(phylum_needed, metadata_with_clinicalparam, by = 'LibraryID')
dim(phylum_data_for_association)
dim(metadata_with_clinicalparam)
dim(phylum_needed)
phylum_list_to_use <-colnames(phylum_needed)[grepl('p__',colnames(phylum_needed))]
```
### GLM
```{r phylumglm}
phylum_with_age <-
  lapply(phylum_data_for_association[, phylum_list_to_use], function(x)
    coefficients(summary(
      glm(
        Age ~ x +  FBGlu +  Tgly + TCHOL + HDL +  BMI + Gender,
        data = phylum_data_for_association))))

phylum_fname <- 'phylum_associations_noldl.csv'
phylum_results <- param_with_age_correction(phylum_list_to_use, phylum_with_age, resultsfoldername, phylum_fname, phylum_data_for_association)

# phylum_maaslin2 <- run_maaslin2(phylum_data_for_association, 'NONE', 'NONE', 'results_0.1')
# phylum_maaslin2_glm <- phylum_maaslin2 %>% 
#   dplyr::select(org_list, pval, coef, stderr, padj) %>% 
#   dplyr::rename(betaestimate = coef, stderror = stderr) %>% 
#   # add 95% confint
#   mutate(`2.5%` = betaestimate - 1.96 * stderror, 
#          `97.5%` = betaestimate + 1.96 * stderror)

```
### Preparing data to plot
```{r }

median_of_phylum <- phylum_needed %>% 
  dplyr::select(rownames(phyluminmorethanfiftysamples)) %>%
  as.matrix() %>%
  matrixStats::colMedians()

median_of_phylum_with_name <- data.frame(rownames(phyluminmorethanfiftysamples), median_of_phylum)

phylum_results$Significant <-
  ifelse(phylum_results$padj < 0.05, "FDR < 0.05", "Not Sig")

colnames(median_of_phylum_with_name)[1] <- c('org_list')

phylum_results_merge <-
  merge(phylum_results, 
        #phylum_maaslin2_glm,
        median_of_phylum_with_name, 
        by = 'org_list')

phylum_results_merge$highlight1 <-
  ifelse(phylum_results_merge$median_of_phylum > 0 &
           phylum_results$padj < 0.05, "2",
         ifelse(phylum_results_merge$median_of_phylum == 0 &
                  phylum_results$padj < 0.05 |phylum_results$padj == 0.05,
                "1", "0"))

phylum_results_merge$org_list <- str_replace_all(phylum_results_merge$org_list, 'p__', '')

phylum_results_merge$org_list <- str_replace_all(phylum_results_merge$org_list, '_', ' ')

write.csv(phylum_results_merge, here(resultsfoldername, phylum_fname))

print(phylum_results_merge)
```
