---
title: "Supplementary Figure 3"
author: "Indrik Wijaya"
date: "2024-02-25"
editor_options: 
  chunk_output_type: console
chunk_output_type: console
output: 
  
  rmdformats::downcute:
    
    downcute_theme: "chaos"
    use_bookdown: true
    self_contained: true
    lightbox: true
    gallery: true
    toc_depth: 5
  pdf_document:
    toc_depth: 5
---

# Load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE, 
                      collapse = TRUE, warning = FALSE, message = FALSE)

library(here)

species_thr <- 0.1
genus_thr <- 0.1
hist_filter <- 50
figsfoldername <- paste0("figs")
dir.create(file.path(here(figsfoldername)))
resultsfoldername <- paste0("results")
dir.create(file.path(here(resultsfoldername)))
editablepptfname <- "figs_ppts"
dir.create(file.path(here(editablepptfname)))
source(here('script', 'utils_DA.R'))
source(here('script', 'utils_prev.R'))
source(here('script','utils_highsens.R'))
source(here('script','fig_plot.R'))
conflicted::conflict_prefer('desc', 'dplyr')
conflicted::conflict_prefer('rename', 'dplyr')
conflicted::conflict_prefer('filter', 'dplyr')

set.seed(100)
```

Run validation tests based on different analyses, normalization and thresholds
- Analyses: GLM on ALL cohorts, GLM on SG90 + CPE only, Prevalence Test (CA) on ALL cohorts
- Methods: GLM, ANCOM-BC, Maaslin2
- Normalization: TSS & CSS 
- % filtering threshold: 0.05%, 0.1%, 0.5%, 1%

# Load data & metadata
```{r}
species_thr <- 0.1
genus_thr <- 0.1
hist_filter <- 50
n_nonzero <- 50

# metadata with imputed values, updated age for SG90
metadata <- read.csv(here('data', 'metadata_with_imputed_updatedage_allInfo.csv'))

# load all profiles with cre & filtered with 0.1% threshold - unadjusted
all_profiles <- read.csv(here('data', 'taxonomic_profiles_filtered.csv')) %>% 
  dplyr::rename(org_list = 'X')

# get SG90 + CPE profiles only
sg90_cpe_metadata <- metadata %>% 
  filter(Cohort == 'SG90') %>% 
  mutate(LibraryID = str_remove(LibraryID, '[.]'))
sg90_cpe_profiles <- all_profiles %>% 
  dplyr::select(org_list, sg90_cpe_metadata$LibraryID) %>% 
  tibble::column_to_rownames('org_list') #382 * 295

# remove all zeros
sg90_cpe_profiles <- sg90_cpe_profiles[rowSums(sg90_cpe_profiles) != 0, ]
sg90_cpe_profiles <- sg90_cpe_profiles %>% 
  tibble::rownames_to_column('org_list') #330 * 296

# remove sg90 samples with no date for blood collection - 18 samples
no_blood_date <- read.csv(here('data', 'no_blood_date.csv'))

# old metadata with no imputed values - to get patient id
old_metadata <- read.csv(here('data', 'metadata_with_clinical_data.csv')) %>% 
  dplyr::rename(LibraryID = 'Index')

# get libraryid - 17 samples, 1 sample doesn't have a libraryID
no_blood_date <- no_blood_date %>% 
  mutate(Patient.ID = as.character(Patient.ID)) %>% 
  left_join(old_metadata, by = 'Patient.ID') %>% 
  dplyr::filter(!is.na(LibraryID))



```

## Convert relative abundance into estimated read count by multiplying through sequencing depth
```{r}
combined_sanity <- read.csv(here('data', 'combined_sanity.csv'))

read_count <- combined_sanity %>% 
  filter(LibraryID %in% names(all_profiles)[-1]) %>% 
  mutate(fq_readcount = fq_readcount/2) %>% 
  dplyr::select(LibraryID, fq_readcount)

# load all profiles with cre & filtered with 0.1% threshold - adjusted
all_profiles_adj <- read.csv(here('data', 'taxonomic_profiles_batch_adjusted.csv')) %>% 
  dplyr::rename(org_list = 'X')

all_profiles_count <- all_profiles_adj
for (lib in read_count$LibraryID) {
  #print(lib)
  seq_depth <- read_count %>% 
    filter(LibraryID == lib) %>% 
    pull(fq_readcount)
  #print(seq_depth)
  all_profiles_count[[lib]] <- all_profiles_count[[lib]] * seq_depth 
}

all_profiles_count_assoc <- transform_tax_profiles(all_profiles_count, 
                                                    metadata, 50)

```

# Original + CPE

## EM
```{r}
all_profiles_assoc <- transform_tax_profiles(all_profiles_adj, metadata, 50)
all_glm <- run_glm(all_profiles_assoc)
```

## ANCOM-BC

```{r, results='hide'}
# run ancom-bc
tax_ps_all <- build_ps(all_profiles_count_assoc)

ancombc_all_count_out <- ancombc(phyloseq = tax_ps_all,
                                  formula = 'Age + Gender + FBGlu + Tgly + TCHOL + HDL + BMI',
                                  p_adj_method = 'fdr', prv_cut = 0, lib_cut = 0,
                                  neg_lb = F, tol = 1e-5,
                                  max_iter = 100, conserve = T, alpha = 0.05)

ancombc_all_da <- get_ancombc_da(ancombc_all_count_out)
names(ancombc_all_da)[1] <- 'org_list'
plot_ancombc_da(ancombc_all_da)

ancombc_all_out_glm <- convert_ancombc_to_glm(ancombc_all_count_out)
# reorder columns
ancombc_all_out_glm <- ancombc_all_out_glm[, c('org_list', 'pval', 'betaestimate', 
                                                 'stderror', 'padj')]

```

## Maaslin2

```{r, results='hide'}
maaslin2_all_res <- run_maaslin2(all_profiles_assoc, 'NONE', 'NONE', 'association_with_age')
maaslin2_all_res_glm <- maaslin2_all_res %>% 
  dplyr::select(org_list, pval, coef, stderr, padj) %>% 
  dplyr::rename(betaestimate = coef, stderror = stderr)

```

## CSS
```{r}
all_profiles_css <- CSSnorm(all_profiles_adj)
all_profiles_css_assoc <- transform_tax_profiles(all_profiles_css, metadata, 50)
all_css_glm <- run_glm(all_profiles_css_assoc)
# all_css_maaslin2 <- run_maaslin2(all_profiles_assoc, 'NONE', 'NONE', 
#                                  'association_with_age_css')
# maaslin2_all_res_css <- all_css_maaslin2 %>% 
#   dplyr::select(org_list, pval, coef, stderr, padj) %>% 
#   dplyr::rename(betaestimate = coef, stderror = stderr)

```

# SG90 + CPE only

## EM
```{r}
sg90_cpe_association <- transform_tax_profiles(sg90_cpe_profiles, sg90_cpe_metadata, 50)

sg90_cpe_glm <- run_glm(sg90_cpe_association)

```

## Get count
```{r}
read_count_sg90_cpe <- combined_sanity %>% 
  filter(LibraryID %in% names(sg90_cpe_profiles)[-1]) %>% 
  mutate(fq_readcount = fq_readcount/2) %>% 
  dplyr::select(LibraryID, fq_readcount)

sg90_cpe_profiles_count <- sg90_cpe_profiles
for (lib in read_count_sg90_cpe$LibraryID) {
  #print(lib)
  seq_depth <- read_count_sg90_cpe %>% 
    filter(LibraryID == lib) %>% 
    pull(fq_readcount)
  #print(seq_depth)
  sg90_cpe_profiles_count[[lib]] <- sg90_cpe_profiles_count[[lib]] * seq_depth 
}

sg90_cpe_association_count <- transform_tax_profiles(sg90_cpe_profiles_count, 
                                                    sg90_cpe_metadata, 50)

```

## ANCOM-BC
```{r, results='hide'}
# run ancom-bc
tax_ps_sg90_cpe <- build_ps(sg90_cpe_association_count)

ancombc_sg90_cpe_count_out <- ancombc(phyloseq = tax_ps_sg90_cpe,
                                  formula = 'Age + Gender + FBGlu + Tgly + TCHOL + HDL + BMI',
                                  p_adj_method = 'fdr', prv_cut = 0, lib_cut = 0,
                                  neg_lb = F, tol = 1e-5,
                                  max_iter = 100, conserve = T, alpha = 0.05)

ancombc_sg90_cpe_da <- get_ancombc_da(ancombc_sg90_cpe_count_out)
names(ancombc_sg90_cpe_da)[1] <- 'org_list'
plot_ancombc_da(ancombc_sg90_cpe_da)

ancombc_sg90_cpe_out_glm <- convert_ancombc_to_glm(ancombc_sg90_cpe_count_out)
# reorder columns
ancombc_sg90_cpe_out_glm <- ancombc_sg90_cpe_out_glm[, c('org_list', 'pval', 'betaestimate', 'stderror', 'padj')]
```

## Maaslin2

```{r, results='hide'}
maaslin2_sg90_cpe_res <- run_maaslin2(sg90_cpe_association, 'NONE', 'NONE', 'association_with_age')
maaslin2_sg90_cpe_res_glm <- maaslin2_sg90_cpe_res %>% 
  dplyr::select(org_list, pval, coef, stderr, padj) %>% 
  dplyr::rename(betaestimate = coef, stderror = stderr)
```

## CSS
```{r}
# filter zero species
sg90_cpe_css <- CSSnorm(sg90_cpe_profiles)
sg90_cpe_css_assoc <- transform_tax_profiles(sg90_cpe_css, sg90_cpe_metadata, 50)

sg90_cpe_css_glm <- run_glm(sg90_cpe_css_assoc)

# sg90_cpe_css_maaslin2 <- run_maaslin2(sg90_cpe_association, 'NONE', 'NONE', 
#                                       'association_with_age_css')
# maaslin2_sg90_cpe_res_css <- sg90_cpe_css_maaslin2 %>% 
#   dplyr::select(org_list, pval, coef, stderr, padj) %>% 
#   dplyr::rename(betaestimate = coef, stderror = stderr)
```

# Prevalence Test - Cochran Armitage

## Apply Different thresholds
```{r, results='hide'}
# load all profiles with cre & filtered with 0.1% threshold - unadjusted
all_profiles_unfilt <- read.csv(here('data', 'taxonomic_profiles.csv')) %>% 
  tibble::column_to_rownames('X') %>% 
  replace(is.na(.), 0) 
## renormalize to 100%
all_profiles_unfilt <- data.frame(apply(all_profiles_unfilt, 2, function(x) x/sum(x)) * 100) %>% 
  tibble::rownames_to_column('org_list')
combined_profiles_filt_0.01 <- remove_low_abund(all_profiles_unfilt, 0.01) #535 taxa
combined_profiles_filt_0.05 <- remove_low_abund(all_profiles_unfilt, 0.05) #434 taxa
combined_profiles_filt_0.1 <- remove_low_abund(all_profiles_unfilt, 0.1) #382 taxa
combined_profiles_filt_0.5 <- remove_low_abund(all_profiles_unfilt, 0.5) #284 taxa
combined_profiles_filt_1 <- remove_low_abund(all_profiles_unfilt, 1) #234 taxa
combined_profiles_filt_2 <- remove_low_abund(all_profiles_unfilt, 2) #186 taxa

```

## Only use taxa that have prevalence in >= 50 samples
- to shrink the list and match according to original list of taxa used
```{r}
## 0.1%
combined_profiles_filt_0.1_association <- 
  transform_tax_profiles(combined_profiles_filt_0.1 %>% 
                           tibble::rownames_to_column('org_list'),
                                                                  metadata, 50)
org_list_to_use_0.1 <- colnames(combined_profiles_filt_0.1_association)[grepl('__', colnames(combined_profiles_filt_0.1_association))]

combined_for_test_0.1 <- get_profiles_for_test(combined_profiles_filt_0.1, org_list_to_use_0.1)

combined_count_age_group_0.1 <- get_count_age_group(combined_for_test_0.1, metadata) #104 taxa

## 0.01%
combined_profiles_filt_0.01_association <- 
  transform_tax_profiles(combined_profiles_filt_0.01 %>% 
                           tibble::rownames_to_column('org_list'),
                                                                  metadata, 50)
org_list_to_use_0.01 <- colnames(combined_profiles_filt_0.01_association)[grepl('__', colnames(combined_profiles_filt_0.01_association))]

combined_for_test_0.01 <- get_profiles_for_test(combined_profiles_filt_0.01, org_list_to_use_0.01)

combined_count_age_group_0.01 <- get_count_age_group(combined_for_test_0.01, metadata) #162 taxa

## 0.05%
combined_profiles_filt_0.05_association <- 
  transform_tax_profiles(combined_profiles_filt_0.05 %>% 
                           tibble::rownames_to_column('org_list'),
                         metadata, 50)
org_list_to_use_0.05 <- 
  colnames(combined_profiles_filt_0.05_association)[grepl('__', colnames(combined_profiles_filt_0.05_association))]

combined_for_test_0.05 <- get_profiles_for_test(combined_profiles_filt_0.05, org_list_to_use_0.05)

combined_count_age_group_0.05 <- get_count_age_group(combined_for_test_0.05, metadata) #128 taxa


## 0.5%
combined_profiles_filt_0.5_association <- transform_tax_profiles(combined_profiles_filt_0.5 %>% 
                                                                    tibble::rownames_to_column('org_list'),
                                                                  metadata, 50)
org_list_to_use_0.5 <- colnames(combined_profiles_filt_0.5_association)[grepl('__', colnames(combined_profiles_filt_0.5_association))]

combined_for_test_0.5 <- get_profiles_for_test(combined_profiles_filt_0.5, org_list_to_use_0.5)

combined_count_age_group_0.5 <- get_count_age_group(combined_for_test_0.5, metadata) #69 taxa

# 1%
combined_profiles_filt_1_association <- transform_tax_profiles(combined_profiles_filt_1 %>% 
                                                                    tibble::rownames_to_column('org_list'),
                                                                  metadata, 50)
org_list_to_use_1 <- colnames(combined_profiles_filt_1_association)[grepl('__', colnames(combined_profiles_filt_1_association))]

combined_for_test_1 <- get_profiles_for_test(combined_profiles_filt_1, org_list_to_use_1)

combined_count_age_group_1 <- get_count_age_group(combined_for_test_1, metadata) #52 taxa

# 2%
combined_profiles_filt_2_association <- transform_tax_profiles(combined_profiles_filt_2 %>% 
                                                                    tibble::rownames_to_column('org_list'),
                                                                  metadata, 50)
org_list_to_use_2 <- colnames(combined_profiles_filt_2_association)[grepl('__', colnames(combined_profiles_filt_2_association))]

combined_for_test_2 <- get_profiles_for_test(combined_profiles_filt_2, org_list_to_use_2)

combined_count_age_group_2 <- get_count_age_group(combined_for_test_2, metadata) #38 taxa

```


## Run CA Test
```{r}
# use all taxa that pass 0.1% threshold
# check whether profiles is in the order of 100% or 1
combined_ca_large_with_dir_0.01 <- run_cohran_armitage(combined_count_age_group_0.01, metadata)

combined_ca_large_with_dir_0.05 <- run_cohran_armitage(combined_count_age_group_0.05, metadata)
combined_ca_large_with_dir_0.1 <- run_cohran_armitage(combined_count_age_group_0.1, metadata)

combined_ca_large_with_dir_0.5 <- run_cohran_armitage(combined_count_age_group_0.5, metadata)
combined_ca_large_with_dir_1 <- run_cohran_armitage(combined_count_age_group_1, metadata)

combined_ca_large_with_dir_2 <- run_cohran_armitage(combined_count_age_group_2, metadata)



```

# Get High Sensitivity Results 
- significant in >= 2 methods and are of the same directions
- take median p-adj values, beta from all cohorts
```{r,warning=FALSE}
# high sensitivity list
high_sens_scores <- get_combined_scores(all_glm, sg90_cpe_glm, combined_ca_large_with_dir_0.1)
high_sens_cons_scores <- get_cons_score(high_sens_scores)
high_sens <- filter_same_dir(high_sens_cons_scores)
high_sens_with_med_padj <- get_median_padj(
  all_glm, sg90_cpe_glm, combined_ca_large_with_dir_0.1)

# save all profiles glm with median padj
all_med_glm <- #maaslin2_all_res_glm %>%
  all_glm %>% 
  dplyr::select(-padj) %>% 
  left_join(high_sens_with_med_padj %>% dplyr::select(-n) %>% 
              dplyr::rename(padj = 'padj_median'),
            by = 'org_list')
# write_tsv(all_med_glm, here('results_0.1', 'species_associations.tsv'))

```

# Combine Results - SuppFig3
```{r,warning=FALSE}
# get final significant species

combined_scores <- rbind(
   
   all_glm %>% mutate(Group = 'All Cohorts\nEM') %>%
      dplyr::select(org_list:padj, Group),
   all_css_glm %>% mutate(Group = 'All Cohorts\nCSS') %>%
   dplyr::select(org_list:padj, Group),
   # maaslin2_all_res_css %>% mutate(Group = 'All Cohorts\nCSS') %>% 
   #   dplyr::select(org_list:padj, Group),
   ancombc_all_out_glm %>% mutate(Group = 'All Cohorts\nANCOM-BC') %>% 
      dplyr::select(org_list:padj, Group),
  maaslin2_all_res_glm %>% mutate(Group = 'All Cohorts\nTSS') %>%
   dplyr::select(org_list:padj, Group),

   # 
   sg90_cpe_glm %>% mutate(Group = 'SG90 + CPE\nEM') %>%
      dplyr::select(org_list:padj, Group),
   sg90_cpe_css_glm %>% mutate(Group = 'SG90 + CPE\nCSS') %>%
      dplyr::select(org_list:padj, Group),
   # maaslin2_sg90_cpe_res_css %>% mutate(Group = 'SG90 + CPE\nCSS') %>%
   #    dplyr::select(org_list:padj, Group),
  ancombc_sg90_cpe_out_glm %>% mutate(Group = 'SG90 + CPE\nANCOM-BC') %>% 
      dplyr::select(org_list:padj, Group),
  maaslin2_sg90_cpe_res_glm %>% mutate(Group = 'SG90 + CPE\nTSS') %>%
      dplyr::select(org_list:padj, Group),
   # 
   combined_ca_large_with_dir_0.05 %>% mutate(Group = 'Prevalence\n0.05%') %>% 
      dplyr::select(org_list:padj, Group),
   combined_ca_large_with_dir_0.1 %>% mutate(Group = 'Prevalence\n0.1%') %>%
      dplyr::select(org_list:padj, Group),
   combined_ca_large_with_dir_0.5 %>% mutate(Group = 'Prevalence\n0.5%') %>%
      dplyr::select(org_list:padj, Group),
   combined_ca_large_with_dir_1 %>% mutate(Group = 'Prevalence\n1%') %>% 
      dplyr::select(org_list:padj, Group)
   # combined_fisher_2 %>% mutate(Group = 'Prevalence\n2%') %>%
   #    dplyr::select(org_list:padj, Group)
   
) 
combined_scores <- combined_scores %>% 
    # 0. remove stderror column
    dplyr::select(-stderror) %>% 
    # 1. filter only significant taxa
    #filter(org_list %in% signif_taxa) %>% 
    # 2. filter pval < 0.05
    filter(pval < 0.05) %>% 
    # 3. make long to plot
    pivot_longer(-c('org_list', 'Group')) %>% 
    # 4. remove space, _ from org_list
    mutate(org_list = str_replace_all(org_list, '_', ' '))


# get no of signif taxa per Group
combined_scores %>% 
  filter(name == 'padj', value < 0.05) %>% 
  dplyr::count(Group)

# get beta only
combined_scores_beta <- combined_scores %>% filter(name == 'betaestimate') %>% 
  pivot_wider(names_from = 'Group', values_from = 'value') %>% 
  dplyr::select(-name) %>% 
  pivot_longer(-org_list, names_to = 'Group') %>% 
  mutate(value = as.factor(sign(value)),
         org_list = factor(org_list, levels = rev(sort(unique(combined_scores$org_list)))),
         # Group = factor(Group, levels = c('Original\nNo SPMP',
         #                                  'Original\nWith Old SPMP',
         #                                  'Original\nFIML',
         #                                  'With CPE\nFIML',
         #                                  'Fisher')))
         Group = factor(Group, levels = unique(combined_scores$Group)))

# get significant taxa only >= 1 method
combined_scores_padj <- combined_scores %>% 
  filter(name == 'padj', value < 0.05) %>% 
  dplyr::select(-value)

## get consistency score
score_table <- get_cons_score(combined_scores)
  
  # filter score >= 0.5
score_filter <- score_table %>% 
  filter(Score >= 0.5)

methods_heatmap <- combined_scores_beta %>% 
    # filter taxa w/ 2 out of 3 from previous analysis
    filter(org_list %in% high_sens$org_list) %>%
    ggplot(aes(x = Group, y = org_list, fill = value)) + 
    geom_tile() + 
    geom_tile(colour = 'black', linewidth = 0.3, show.legend = F, alpha = 0.5) +
    geom_text(data = combined_scores_padj %>% 
                filter(org_list %in% high_sens$org_list),
              aes(x = Group, y = org_list, label = 'X'), size = 4,
              color = 'white', inherit.aes = F) +
    scale_fill_manual(values = c('blue', 'red'), na.value = 'white',
                      labels = c('Negative', 'Positive', NA),
                      na.translate=F) +
    scale_x_discrete(expand = c(0, 0)) + 
    scale_y_discrete(expand = c(0, 0)) +
    theme_bw() + 
    coord_fixed(ratio=0.265) +
    labs(fill = 'Direction') + 
    # scale_x_discrete(labels = c('SG90 + CPE', 'SG90 + CPE', 'SG90 + CPE', 'SG90 + CPE',
    #                             'All Cohorts', 'All Cohorts', 'All Cohorts', 'All Cohorts',
    #                             'All Cohorts', 'All Cohorts', 'All Cohorts', 'All Cohorts')) +
    theme(axis.text.y = element_text(face = 'italic', size = 11,
                                     margin = margin(r=0),
                                     hjust = 1.05),
          axis.text.x = element_blank(),
          axis.title = element_blank(), 
          plot.background = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(), 
          text = element_text(family = 'Arial'), 
          legend.position = 'none', 
          axis.ticks = element_blank(),
          #plot.margin=grid::unit(c(0,0,0,0), "mm"))
          plot.margin=margin(0, 0, 0, 0, 'cm'),
          panel.spacing = unit(0, 'cm'))
  
methods_heatmap

# get wide version
combined_table_sign_mat <- score_table %>% 
  filter(org_list %in% high_sens$org_list) %>% 
    tibble::column_to_rownames('org_list')

# get significant results per method
colSums(combined_table_sign_mat != 0)
```

## Use bar annotation to annotate x axis labels

```{r,warning=FALSE}
sg90_cpe_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[1], alpha=0.5 ))
sg90_cpe_textgrob <- text_grob(x=1, y=1, label='SG90 + CPE', size=10)

with_cpe_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[2], alpha=0.5 ))
with_cpe_textgrob <- text_grob(x=1, y=1, label='All Cohorts', size=10)

tss_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[3], alpha=0.5 ))
tss_textgrob <- text_grob(x=1, y=1, label='TSS', size=10)

css_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[4], alpha=0.5 ))
css_textgrob <- text_grob(x=1, y=1, label='CSS', size=10)

em_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[10], alpha=0.5 ))
em_textgrob <- text_grob(x=1, y=1, label='EM', size=10)

ancombc_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[6], alpha=0.5 ))
ancombc_textgrob <- text_grob(x=1, y=1, label='ANCOM\nBC', size=10)

maaslin2_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[7], alpha=0.5 ))
maaslin2_textgrob <- text_grob(x=1, y=1, label='Maaslin2', size=9)

prevalence_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[8], alpha=0.5 ))
prevalence_textgrob <- text_grob(x=1, y=1, label = 'Prevalence', size=10)

cohort_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[9], alpha=0.5 ))
cohort_textgrob <- text_grob(x=1, y=1, label = 'Cohort', size=10)

norm_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[5], alpha=0.5 ))
norm_textgrob <- text_grob(x=1, y=1, label = 'Normalization', size=10)

analysis_grob <- rectGrob(x=1, y=1, gp=gpar( fill=rainbow(11)[11], alpha=0.5 ))
analysis_textgrob <- text_grob(x=1, y=1, label = 'Analysis\ Methods', size=10)

thresh0.05_textgrob <- text_grob(x=1, y=1, label = '0.05%', size=10)
thresh0.1_textgrob <- text_grob(x=1, y=1, label = '0.1%', size=10)
thresh0.5_textgrob <- text_grob(x=1, y=1, label = '0.5%', size=10)
thresh1_textgrob <- text_grob(x=1, y=1, label = '1%', size=10)
thresh2_textgrob <- text_grob(x=1, y=1, label = '2%', size=10)

methods_heatmap_box <- methods_heatmap + 
  coord_cartesian(y = range(seq_along(rownames(combined_table_sign_mat))), clip = 'off') + 
  theme(plot.margin = unit(c(0.015, 0, 0.1, 0), units="npc")) +
  
  # 1st layer - Identifiers
  #annotation_custom(grob = cohort_grob, xmin=-2.6, xmax=-0.55, ymin=-1.5, ymax=-0.3) +
  annotation_custom(grob = cohort_textgrob, xmin=-1, xmax=0, ymin=-1.5, ymax=-0.3) +

  # 1st layer - Cohorts
  annotation_custom(grob = with_cpe_grob, xmin=-1.5, xmax=2.5, ymin=-1.5, ymax=-0.3) +
  annotation_custom(grob = with_cpe_textgrob, xmin=-1.5, xmax=2.5, ymin=-1.5, ymax=-0.3) +

  annotation_custom(grob = sg90_cpe_grob, xmin=2.5, xmax=6.5, ymin=-1.5, ymax=-0.3) +
  annotation_custom(grob = sg90_cpe_textgrob, xmin=2.5, xmax=6.5, ymin=-1.5, ymax=-0.3) +
  
  annotation_custom(grob = with_cpe_grob, xmin=6.5, xmax=10.5, ymin=-1.5, ymax=-0.3) +
  annotation_custom(grob = with_cpe_textgrob, xmin=6.5, xmax=10.5, ymin=-1.5, ymax=-0.3) +

  # 2nd layer - Norm
  #annotation_custom(grob = norm_grob, xmin=-2.6, xmax=-0.55, ymin=-2.8, ymax=-1.6) +
  annotation_custom(grob = norm_textgrob, xmin=-1, xmax=0, ymin=-2.8, ymax=-1.6) +

  # 2nd layer - SG90 + CPE
  annotation_custom(grob = tss_grob, xmin = 0, xmax = 1, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = tss_textgrob, xmin = 0, xmax = 1, ymin = -2.8, ymax = -1.6) +
  
  annotation_custom(grob = css_grob, xmin = 1, xmax = 2, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = css_textgrob, xmin = 1, xmax = 2, ymin = -2.8, ymax = -1.6) +
  
  annotation_custom(grob = ancombc_grob, xmin = 2, xmax = 3, ymin = -4.75, ymax = -2.25) +
  annotation_custom(grob = ancombc_textgrob, xmin = 2, xmax = 3, ymin = -4.75, ymax = -2.25) +
  
  annotation_custom(grob = tss_grob, xmin = 3, xmax = 4, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = tss_textgrob, xmin = 3, xmax = 4, ymin = -2.8, ymax = -1.6) +
  
  # 2nd layer - All cohorts
  annotation_custom(grob = tss_grob, xmin = 4, xmax = 5, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = tss_textgrob, xmin = 4, xmax = 5, ymin = -2.8, ymax = -1.6) +
  
  annotation_custom(grob = css_grob, xmin = 5, xmax = 6, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = css_textgrob, xmin = 5, xmax = 6, ymin = -2.8, ymax = -1.6) +
  
  annotation_custom(grob = ancombc_grob, xmin = 6, xmax = 7, ymin = -4.75, ymax = -2.25) +
  annotation_custom(grob = ancombc_textgrob, xmin = 6, xmax = 7, ymin = -4.75, ymax = -2.25) +
  
  annotation_custom(grob = tss_grob, xmin = 7, xmax = 8, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = tss_textgrob, xmin = 7, xmax = 8, ymin = -2.8, ymax = -1.6) +
  
  # 2nd layer - thresholds
  annotation_custom(grob = thresh0.05_textgrob, xmin = 8, xmax = 9, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = thresh0.1_textgrob, xmin = 9, xmax = 10, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = thresh0.5_textgrob, xmin = 10, xmax = 11, ymin = -2.8, ymax = -1.6) +
  annotation_custom(grob = thresh1_textgrob, xmin = 11, xmax = 12, ymin = -2.8, ymax = -1.6) +
  #annotation_custom(grob = thresh2_textgrob, xmin = 11, xmax = 12, ymin = -2.8, ymax = -1.6) +
  
  # 3rd layer - Analysis
  #annotation_custom(grob = analysis_grob, xmin=-2.6, xmax=-0.55, ymin=-4.1, ymax=-2.9) +
  annotation_custom(grob = analysis_textgrob, xmin=-1.1, xmax=-0.1, ymin=-4.1, ymax=-2.9) +

  # 3rd layer - SG90 + CPE
  annotation_custom(grob = em_grob, xmin = 0, xmax = 1, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = em_textgrob, xmin = 0, xmax = 1, ymin = -4.1, ymax = -2.9) +
  
  annotation_custom(grob = em_grob, xmin = 1, xmax = 2, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = em_textgrob, xmin = 1, xmax = 2, ymin = -4.1, ymax = -2.9) +
  
  # annotation_custom(grob = ancombc_grob, xmin = 2, xmax = 3, ymin = -4.1, ymax = -2.9) +
  # annotation_custom(grob = ancombc_textgrob, xmin = 2, xmax = 3, ymin = -4.1, ymax = -2.9) +
  
  annotation_custom(grob = maaslin2_grob, xmin = 3, xmax = 4, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = maaslin2_textgrob, xmin = 3, xmax = 4, ymin = -4.1, ymax = -2.9) +
  
  # 3rd layer - All cohorts
  annotation_custom(grob = em_grob, xmin = 4, xmax = 5, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = em_textgrob, xmin = 4, xmax = 5, ymin = -4.1, ymax = -2.9) +
  
  annotation_custom(grob = em_grob, xmin = 5, xmax = 6, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = em_textgrob, xmin = 5, xmax = 6, ymin = -4.1, ymax = -2.9) +
  
  # annotation_custom(grob = ancombc_grob, xmin = 6, xmax = 7, ymin = -4.1, ymax = -2.9) +
  # annotation_custom(grob = ancombc_textgrob, xmin = 6, xmax = 7, ymin = -4.1, ymax = -2.9) +
  
  annotation_custom(grob = maaslin2_grob, xmin = 7, xmax = 8, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = maaslin2_textgrob, xmin = 7, xmax = 8, ymin = -4.1, ymax = -2.9) +
  
  # 3rd layer - Prevalence
  annotation_custom(grob = prevalence_grob, xmin = 6.5, xmax = 10.5, ymin = -4.1, ymax = -2.9) +
  annotation_custom(grob = prevalence_textgrob, xmin = 6.5, xmax = 10.5, ymin = -4.1, ymax = -2.9)
  
  
methods_heatmap_box


```

## Add score
```{r, warning=FALSE}
score_heatmap <- score_table %>% 
  filter(org_list %in% high_sens$org_list) %>% 
    pivot_longer(-org_list) %>% 
    filter(name == 'Score') %>% 
    mutate(#org_list = factor(org_list, levels = rev(rownames(combined_table_sign_mat))),
          org_list = factor(org_list, levels = rev(sort(high_sens$org_list))),   
          value = round(value, 2),
           ) %>% 
    ggplot(aes(x = name, y = org_list)) + 
    geom_tile(fill = NA, na.value = NA) + geom_text(aes(label = value), size=3.5,
                                          family = "Arial") +
    scale_x_discrete(expand = c(0, 0)) + 
    scale_y_discrete(expand = c(0, 0)) +
    theme_bw() + 
    theme(axis.text.y = element_blank(),
          axis.text.x = element_text(size = 10),
          axis.title = element_blank(), 
          plot.background = element_blank(), 
          panel.border = element_blank(), 
          panel.grid = element_blank(), 
          text = element_text(family = 'Arial'), 
          legend.position = 'none', 
          axis.ticks = element_blank(), 
          plot.margin = unit(c(0, 0, 0.07, 0), units="npc"))

p <- ggarrange(methods_heatmap_box, NULL, score_heatmap,
          widths = c(13, -0.05, 0.8), nrow=1)
print(p)

template_to_save(p, fname = "Supp_Fig3.png", height = 18, width=24) 

```

## Save editable ppt
```{r, results='hide',warning=FALSE}
suppfig3_dml <- rvg::dml(ggobj = p)
officer::read_pptx() %>%
  # add slide ----
  officer::add_slide() %>%
  # specify object and location of object ----
  officer::ph_with(suppfig3_dml, officer::ph_location(width = 10, height = 7)) %>%
  # export slide -----
  base::print(target = paste0(editablepptfname, "Supp_Fig3.pptx"))
```

## Extras
### QC to check difference in count due to thresholds
```{r}
#test <- rbind(combined_count_age_group_0.1 %>% pivot_longer(-org_list) %>% 
              #     mutate(thresh = 'A'), 
              # combined_count_age_group_2 %>% #pivot_longer(-org_list) %>% 
#                  mutate(thresh = 'B'))
# test %>% 
#   pivot_wider(names_from= 'thresh', values_from= 'value') %>% 
#     ggplot(aes(x = A, y = B)) + 
#     geom_point() + 
#   labs(x = 'Thresh = 0.1%', y = 'Thresh = 2%') +
#     facet_wrap(~name, scales = 'free') + 
#   theme_bw() + 
#   theme(text = element_text(size = 18))

#plot_desc(combined_count_age_group_0.1 %>% 
#            filter(org_list %in% (combined_ca_large_with_dir_0.1 %>% 
#                     filter(padj < 0.05) %>% 
#                     pull(org_list))), metadata)
```

### Intersection of results
```{r}
# Reduce(intersect, list(combined_ca_large_with_dir_0.01 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.05 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.1 %>% filter(padj < 0.05) %>% pull(org_list)))
# 
# Reduce(intersect, list(combined_ca_large_with_dir_0.01 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.05 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.1 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_0.5 %>% filter(padj < 0.05) %>% pull(org_list)))

# Reduce(intersect, list(combined_ca_large_with_dir_0.01 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.05 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.1 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_0.5 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_1 %>% filter(padj < 0.05) %>% pull(org_list)))

# Reduce(intersect, list(combined_ca_large_with_dir_0.01 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.05 %>% filter(padj < 0.05) %>% pull(org_list), 
#           combined_ca_large_with_dir_0.1 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_0.5 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_1 %>% filter(padj < 0.05) %>% pull(org_list),
#        combined_ca_large_with_dir_2 %>% filter(padj < 0.05) %>% pull(org_list)))

```

